### 2020-04-25 ms_8.Rmd
Commit some models (app probably not running now)

---
title: "Maps"
author: "Julia Englebert"
date: "4/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(scipen = 999)

library(shiny)
library(haven)
library(tidyverse)
library(gt)
library(devtools)
library(tibble)
library(foreign)
library(scales)
library(shinythemes)
library(plotly)
library(leaflet)
library(broom)
library(readxl)

# packages for mapping
library(leaflet)
library(spdplyr)
library(rmapshaper)
library(sf)
```

```{r testModel}
#
# model1 <- lm(data = mwdata, perc_school ~ state + land_in_farms)
```

# DATA WRANGLING

```{r allData, cache=TRUE}
# 1800s data

# Will comment on 1870s data, then repeat the process for all subsequent years
# Note that the codenames ARE DIFFERENT FOR EACH YEAR
# Even if the variable was measured in the same way
# See original codebook for details

# Read in all 1870 csv files
# Originally, R was returning this message:
# Column `STATE` joining factors with different levels, coercing to character vectorColumn `COUNTY` joining
# factors with different levels, coercing to character vector
# I got rid of it by converting all factors to characters

data1870a <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds16_1870_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1870b <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds17_1870_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

# Join the two datasets by STATEA and COUNTYA, which are the state and county codes

data1870j <- full_join(data1870a, data1870b, by = c("STATEA", "COUNTYA")) %>%

  # Mutate to get rid of the .x suffix on variables

  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%

  # Mutate to rename varialbes of interest to uniform, comprehensible names
  # Some existing varialbes are combined to create the desired variables
  # See codebook

  mutate(value_farm = AJV001, n_farm = AJS001, cows =  AJ0003+AJ0005, horses = AJ0001, 
         avg_size = (AJT001*2 + AJT002*6 + AJT003*15 + AJT004*35 + AJT005*75 + AJT006*300 + AJT007*750 + 
                       AJT008*1200)/AKP001, value_implements = AKL001, pop = AJ3001, male = AKD001, 
         female = AKD002, enrolled = ALG001, urban_pop = AKE001, pct_urban = AKE001/AJ3001*100, 
         illiterate = AJ7001, acres_farms = AJU001+ AJU002+AJU003,

         # Even though irrigation data doesn't exist for this year, I'm filling the columns in with NA
         # Later years will have irrigation data

         acres_irrigated = NA, pct_irrigated = NA) %>%

  # Mutate again to make new variables out of the ones you just created in the last mutate

  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%

  # Select the desired variables

  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm,
         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female,
         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)

# Now the dataset is much more manageable!

# Start on the next year, 1880, repeating the exact same process
# Lots of NA values due to scant data this year

data1880a <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds22_1880_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1880b <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds23_1880_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1880j <- full_join(data1880a, data1880b, by = c("STATEA", "COUNTYA")) %>%
  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
  mutate(value_farm = APN001+APO001+APQ001, n_farm = AOQ001, cows =  AOG004+AOG005, horses = AOG001, 
         avg_size = APG001, value_implements = APO001, pop = AOT001, male = AP1001, female = AP1002, 
         enrolled = NA, urban_pop = AO4001, pct_urban = AO4001/AOT001*100, illiterate = NA, 
         acres_farms = AOR001, acres_irrigated = NA, pct_irrigated = NA) %>%
  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)

# Start on 1890 data

data1890a <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds26_1890_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1890b <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds27_1890_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1890j <- full_join(data1890a, data1890b, by = c("STATEA", "COUNTYA")) %>%
  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
  mutate(value_farm = AUK001+AUK002+AUK003, n_farm = AS9001, cows = ASY002, horses = ASY003, 
         avg_size = ATW001, value_implements = AUK002, pop = AUM001, male = AV0001+AV0003+AV0005+AV0007, 
         female = AV0002+AV0004+AV0006+AV0008, enrolled = AVK005+AVK006+AVK007+AVK008, urban_pop = AUU001, 
         pct_urban = AUU001/AUM001*100, illiterate = NA, acres_farms = AUJ001+AUJ002, 
         acres_irrigated = AT2001, pct_irrigated = AT2001/(AUJ001+AUJ002)*100) %>%
  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)

# Before I mutated all of the data, running this code showed that all .x values are the same 
# as all .y values, so it was safe to eliminate the double columns as long as "problems" returned 0
# And it did!

#problems <- bind_rows(data1870j, data1880j, data1890j) %>% 
  #select(STATEA, COUNTYA, GISJOIN.x, GISJOIN.y, YEAR.x, YEAR.y, STATE.x, STATE.y, 
         #COUNTY.x, COUNTY.y, AREANAME.x, AREANAME.y) %>%
  #mutate(same = ifelse(GISJOIN.x == GISJOIN.y & YEAR.x == YEAR.y & STATE.x == STATE.y & 
                         #COUNTY.x == COUNTY.y & AREANAME.x == AREANAME.y, TRUE, FALSE)) %>%
  #filter(same == "FALSE") %>%
  #nrow()

# Read in the 1900s data

# Only do decennial census data
# Intermittent years only have AF12001: estimated population

# 1900

data1900a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds30_1900_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1900b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds31_1900_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1900j <- full_join(data1900a, data1900b, by = c("STATEA", "COUNTYA")) %>%
  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
  mutate(value_farm = AWV001, n_farm = AW3001, cows = AW7001, horses = AW7002, 
         avg_size = AXE001, value_implements = AWW003, pop = AWS001, male = AZ0001,female = AZ0001, 
         enrolled = NA, urban_pop = NA, pct_urban = NA, illiterate = AYS001+AYS002+AYS003+AYS004+AYS005, 
         acres_farms = AWT001, acres_irrigated = AX8001, pct_irrigated = AX8001/AWT001*100) %>%
  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)

# 1910

data1910a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds36_1910_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1910b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds37_1910_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1910j <- full_join(data1910a, data1910b, by = c("STATEA", "COUNTYA")) %>%
  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
  mutate(value_farm = A0F001, n_farm = A1H001, cows = A1F001+A1F002, horses = NA, 
         avg_size = A21001/A1H001, value_implements = A0Q005, pop = A0E001, male = NA, female = NA, 
         enrolled = A4F001, urban_pop = A36001, pct_urban = A36001/A0E001*100, illiterate = A38001, 
         acres_farms = A21001, acres_irrigated = A0G001, pct_irrigated = A0G001/A21001*100) %>%
  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)

#1920

data1920a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds43_1920_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME =  as.character(AREANAME))

data1920b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds224_1920_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1920c <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds210_1920_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1920j <- full_join(data1920a, data1920b, by = c("STATEA", "COUNTYA"))
data1920j <- full_join(data1920j, data1920c, by = c("STATEA", "COUNTYA"))
data1920j <- data1920j %>%
  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
  mutate(value_farm = A8P001+A8P002+A8P003+A8P004, n_farm = AB3P001, cows = NA, horses = NA, 
         avg_size = AB3T001/AB3P001, value_implements = A8P003, pop = A81001, male = A8B001, 
         female = A8B002, enrolled = A7N001+A7N002+A7N003+A7N004, urban_pop = A80001, 
         pct_urban = A80001/A81001*100, illiterate = A7U001, acres_farms = AB3T001, 
         acres_irrigated = AB6L001, pct_irrigated = AB6L002/AB3T001*100) %>%
  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)

#1930
# 1930 has more datasets available than previous years

data1930a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds212_1930_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1930b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds224_1930_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1930c <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds52_1930_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1930d <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds53_1930_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1930e <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds54_1930_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1930f <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds55_1930_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1930j <- full_join(data1930a, data1930b, by = c("STATEA", "COUNTYA"))
data1930j <- full_join(data1930j, data1930c, by = c("STATEA", "COUNTYA"))
data1930j <- full_join(data1930j, data1930d, by = c("STATEA", "COUNTYA"))
data1930j <- full_join(data1930j, data1930e, by = c("STATEA", "COUNTYA"))
data1930j <- full_join(data1930j, data1930f, by = c("STATEA", "COUNTYA"))

data1930j <- data1930j %>%
  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
  mutate(value_farm = BFD001+BFD002+BFD004+ACEO001, n_farm = AB95001, cows = NA, horses = NA, 
         avg_size = ACAA001, value_implements = BFD004, pop = BCX001, male = BC3001, 
         female = BC3002, enrolled = BDT001+BDT002+BDT003+BDT004, urban_pop = BDX001, 
         pct_urban = BDX001/BDP001*100, illiterate = BDV001, acres_farms = AB98001, 
         acres_irrigated = ACEX002, pct_irrigated = ACEX002/AB98001*100) %>%
  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)

# 1940

data1940a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds224_1940_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1940b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds77_1940_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1940c <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds78_1940_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1940j <- full_join(data1940a, data1940b, by = c("STATEA", "COUNTYA"))
data1940j <- full_join(data1940j, data1940c, by = c("STATEA", "COUNTYA"))
data1940j <- data1940j %>%
  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%

  # value_farm is missing the land & livestock components!!
  # this variable only includings buildings and implements

  mutate(value_farm = BWD001+BWD002, n_farm = BY1001, cows = NA, horses = NA, 
         avg_size = BY0001, value_implements = BWD002, pop = BVU001, male = BVY001, 
         female = BVY002, enrolled = BWU001+BWU002+BWU003+BWU004+BWU005, urban_pop = NA, 
         pct_urban = NA, illiterate = NA, acres_farms = BY3001, 
         acres_irrigated = NA, pct_irrigated = NA) %>%
  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, 
         pct_illiterate = illiterate/pop*100) %>%
  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)

# 1950

data1950a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds83_1950_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME =  as.character(AREANAME))

data1950b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds84_1950_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1950j <- full_join(data1950a, data1950b, by = c("STATEA", "COUNTYA")) %>%
  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x, 
         B18001 = as.numeric(B18001)) %>%

  # value_farm is missing the implements & livestock components!!
  # This variable only includings buildings and land

  mutate(value_farm = B3M001, n_farm = B3Z001, cows = NA, horses = NA, 
         avg_size = B3L001, value_implements = NA, pop = B13001, male = B3E001, 
         female = B3E002, enrolled = B2A001+B2A002+B2A003+B2A004+B2A005, urban_pop = B2J001, 
         pct_urban = B2J001/B18001*101, illiterate = NA, acres_farms = B3K001, 
         acres_irrigated = NA, pct_irrigated = NA) %>%
  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)

#1960
# The data are SPARSE yikes

data1960a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds89_1960_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME =  as.character(AREANAME))

data1960b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds90_1960_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1960c <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds91_1960_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1960j <- full_join(data1960a, data1960b, by = c("STATEA", "COUNTYA"))
data1960j <- full_join(data1960j, data1960c, by = c("STATEA", "COUNTYA"))
data1960j <- data1960j %>%
  mutate(value_farm = NA, n_farm = NA, cows = NA, horses = NA, 
         avg_size = NA, value_implements = NA, pop = B47001, male = B5H001, 
         female = B5H002, enrolled = NA, urban_pop = NA, pct_urban = NA, illiterate = NA, 
         acres_farms = NA, acres_irrigated = NA, pct_irrigated = NA) %>%
  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)

```

## Join the cleaned sets

```{r boundData, cache=TRUE}
# Running this code (before I renamed variables) showed that all .x values are the same as all .y values, 
# so it was safe to eliminate the double columns
# as long as problems returned 0 (and it did!)
# this didn't check for .x.x or .y.x. columns, etc.

#problems <- bind_rows(data1900j, data1910j, data1920j, data1930j, data1940j, data1950j, data1960j) %>% 
  #select(STATEA, COUNTYA, GISJOIN.x, GISJOIN.y, YEAR.x, YEAR.y, STATE.x, STATE.y, 
         #COUNTY.x, COUNTY.y) %>%
  #mutate(same = ifelse(GISJOIN.x == GISJOIN.y & YEAR.x == YEAR.y & STATE.x == STATE.y & 
                         #COUNTY.x == COUNTY.y, TRUE, FALSE)) %>%
  #filter(same == "FALSE") %>%
  #nrow()

# Now bind all of the years together!
# Add pct_male and avg_value_farm variables
# Filter for only midwestern states/territories

allData_inflated <- as_tibble(bind_rows(data1870j, data1880j, data1890j, 
                       data1900j, data1910j, data1920j, data1930j, data1940j, data1950j, data1960j)) %>%
  mutate(pct_male = male/pop*100, avg_value_farm = value_farm/n_farm, cows_per_farm = cows_per_person, 
         avg_value_implements = value_implements/n_farm) %>%

  # I mislabeled cows_per_farm as cows_per_person, so I'm doing damage control now

  select(!cows_per_person) %>%
  filter(STATE == "North Dakota" | STATE == "South Dakota" | STATE == "Dakota Territory" | 
           STATE == "Nebraska" | STATE == "Minnesota" | STATE == "Iowa" | STATE == "Missouri" | 
           STATE == "Wisconsin" | STATE == "Illinois" | STATE == "Kansas" | STATE == "Michigan" | 
           STATE == "Indiana" | STATE == "Ohio",

         # Get rid of the weird infinite values in this column

         ! is.infinite(cows_per_farm))

# Now adjust everything for inflation
# I made a spreadsheet based on this website: 
# https://www.officialdata.org/us/inflation/1960?amount=1

inflation <- read_excel("inflation.xlsx")

clean_data <- left_join(allData_inflated, inflation, by = c("YEAR" = "year")) %>%
  # Use the multiplier column to adjust for inflation
  mutate(value_farm = value_farm*multiplier, value_implements = value_implements*multiplier) %>%
  mutate(avg_value_farm = value_farm/n_farm, avg_value_implements = value_implements/n_farm)

# Now save as an RDS file to save on space

saveRDS(clean_data, file = "final/my_data.rds")
saveRDS(clean_data, file = "my_data.rds")

```

```{r mapDataset}

# Now just read in the RDS file
# This is all you need to copy into the shiny app

allData <- readRDS(file = "my_data.rds")

# this is workable data, where NA columns have been removed
# There was too much variable variation & inconsistencies to draw conclusions from all years

subData <- allData %>% 
  filter(YEAR == "1870" | YEAR == "1890" | YEAR == "1910") %>%
  select(-acres_irrigated, -pct_irrigated, -illiterate, -pct_illiterate, -male, -female, -pct_male)

# This next bitwill be used for mapping only!
# All relevant columns must be numeric in order to function properly
# Renaming them also helps make the maps more easily interpretable
# But it's not ideal for working with the data in other contexts, hence the separate set

mwData <- allData %>%
  mutate("Value of Farms" = as.numeric(value_farm), "Average Farm Value" = as.numeric(avg_value_farm), 
         "Value of Implements" = as.numeric(value_implements), 
         "Average Implement Value" = as.numeric(avg_value_implements), 
         "Average Farm Size" = as.numeric(avg_size), "Number of Farms" = as.numeric(n_farm), 
         "Acres of Farmland" = as.numeric(acres_farms), "Irrigated Acres" = as.numeric(acres_irrigated), 
         "Percentage of Farmland Irrigated" = as.numeric(pct_irrigated), "Number of Cows" = as.numeric(cows),
         "Number of Cows per Farm" = as.numeric(cows_per_farm), "Number of Horses" = as.numeric(horses), 
         "Total Population" = as.numeric(pop), "Male Population" = as.numeric(male), 
         "Female Population" = as.numeric(female), "Male Percentage of Population" = as.numeric(pct_male),
         "Total School Enrollment" = as.numeric(enrolled), 
         "Percentage School Enrollment" = as.numeric(pct_enrolled), 
         "Urban Population" = as.numeric(urban_pop), 
         "Percent of Population in Urban Areas" = as.numeric(pct_urban), 
         "Illiterate Population" = as.numeric(illiterate), 
         "Illiteracy Rate" = as.numeric(pct_illiterate)) %>%
  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, 
         "Value of Farms", "Average Farm Value", "Value of Implements", "Average Implement Value",
                   "Average Farm Size", "Number of Farms", "Acres of Farmland", "Irrigated Acres", 
                   "Percentage of Farmland Irrigated", "Number of Cows", "Number of Cows per Farm", 
                   "Number of Horses", "Total Population", "Male Population", "Female Population", 
                   "Male Percentage of Population", "Total School Enrollment", 
                   "Percentage School Enrollment", "Urban Population", 
                   "Percent of Population in Urban Areas", "Illiterate Population", "Illiteracy Rate")

# saving this for later
# will need in shiny app
#sliderOptions <- c("Value of Farms", "Average Farm Value", "Value of Implements", "Average Implement Value",
                   #"Average Farm Size", "Number of Farms", "Acres of Farmland", "Irrigated Acres", 
                   #"Percentage of Farmland Irrigated", "Number of Cows", "Number of Cows per Farm", 
                   #"Number of Horses", "Total Population", "Male Population", "Female Population", 
                   #"Male Percentage of Population", "Total School Enrollment", 
                   #"Percentage School Enrollment", "Urban Population", 
                   #"Percent of Population in Urban Areas", "Illiterate Population", "Illiteracy Rate")

```

# Shiny 
## Preparation

```{r shapefiles}

# Now save the gis data with short names for sake of the shiny app
# hopefully this will work...

dsn1870 <- "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1870"
  layer1870 <- "US_county_1870_conflated"
dsn1880 <- "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1880"
  layer1880 <- "US_county_1880_conflated"
dsn1890 <- "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1890"
  layer1890 <- "US_county_1890_conflated"
dsn1900 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1900"
  layer1900 <- "US_county_1900_conflated"
dsn1910 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1910"
  layer1910 <- "US_county_1910_conflated"
dsn1920 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1920"
  layer1920 <- "US_county_1920_conflated"
dsn1930 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1930"
  layer1930 <- "US_county_1930_conflated"
dsn1940 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1940"
  layer1940 <- "US_county_1940_conflated"
dsn1950 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1950"
  layer1950 <- "US_county_1950_conflated"
dsn1960 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1960"
  layer1960 <- "US_county_1960_conflated"

# Make a tibble of all the shapefiles that you'll need
  shapefiles <- tibble(Year = c(1870, 1880, 1890, 1900, 1910, 1920, 1930, 1940, 1950, 1960), 
                     Dsn = c(dsn1870, dsn1880, dsn1890, dsn1900, dsn1910, dsn1920, dsn1930, 
                             dsn1940, dsn1950, dsn1960),
                     Layer = c(layer1870, layer1880, layer1890, layer1900, layer1910, layer1920, 
                     layer1930, layer1940, layer1950, layer1960))
```

## Output
### Final
#### Change

```{r gisTemplate}
# might also want to add a filter for state??
# if time permits...

# gis template for any year/variable desired
# only change these two lines!
year <- 1870
myvar <- "Average Farm Value"

#choices for myvar will be determined by: 
appPlan <- allData %>%
  filter(YEAR == "1870") %>%
  select_if(~ !all(is.na(.))) %>%
  select(-STATEA:-COUNTY) %>%
  select(-multiplier)

datayear <- mwData %>%
  # use get() to use the stored variable
  # idk why, but it doesn't work without this
  # the variable must be converted to a double in the original mwData dataset
  mutate(myvariable = (get(myvar))) %>%
  filter(YEAR == year, myvariable > 0) %>%
  select(GISJOIN, myvariable)

# Import Census Tract Shapefile into R as SpatialPolygonsDataFrameFormat (SP Dataframe)
# dsn is location of folder which contains shapefiles, (.proj, .shp etc.)
# layer is the filename of the .shp file inside the
# folder dsn points to.

dsnyear <- shapefiles %>% 
  filter(Year == year) %>% 
  pull(Dsn)
layeryear <- shapefiles %>% 
  filter(Year == year) %>% 
  pull(Layer)

countyyear <- sf::st_read(dsn = dsnyear,
                         layer = layeryear)

countyyear <-
  countyyear %>%
  merge(datayear, "GISJOIN")

# Set projection of tracts dataset to `projection` required by leaflet

countyyear <- sf::st_transform(countyyear, crs="+init=epsg:4326")

# Condense size of data for faster processing

countyyear <- rmapshaper::ms_simplify(countyyear)

# Set palette color

pal <- colorNumeric("viridis", NULL)

#  Plot the data

# use shiny to add a title, which will be be equivalent to myvar

leaflet(countyyear) %>%
  addTiles() %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
              fillColor = ~pal(myvariable)) %>%
  addLegend(pal = pal, values = ~myvariable, opacity = 1.0, title = myvar)

```

```{r experimentalModels}
# USE subData FOR ALL MODELS
# EVERYTHING ELSE IS DUBIOUS

# Incorporate
# Value of implements functions as a stand in for mechanization
impedu_plot <- allData %>%
  mutate(YEAR = as.factor(YEAR)) %>%
  group_by(YEAR) %>%
  select(YEAR, avg_value_implements, pct_enrolled) %>%
  na.omit() %>%
  ggplot(aes(x = avg_value_implements, y = pct_enrolled, color = YEAR)) + 
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm", se = FALSE) +
  ylim(0, 45) + 
  labs(title = "Relationship Between Farm Value and Education, by County",
       subtitle = "Midwestern United States", 
       x = "Average Value of Implements", 
       y = "Percent Population Enrolled in School") +
  scale_x_log10(labels = comma) +
  theme_classic()

# Incorporate
# States are gaining more implements each year
imp_plot <- allData %>%
  mutate(YEAR = as.factor(YEAR)) %>%
  select(YEAR, STATE, value_implements) %>%
  na.omit() %>%
  group_by(YEAR, STATE) %>%
  summarize(mean_imp = mean(value_implements)) %>% 
  ungroup() %>%
  group_by(STATE) %>%
  ggplot(aes(x = YEAR, y = mean_imp, fill = STATE)) + 
  geom_col() +
  labs(title = "Implement Value by Year, by State",
       subtitle = "Dollar Values Adjusted for Inflation",
       x = "Year",
       y = "Average Value of All Implements",
       fill = "State") +
  theme_classic()

# INCORPORATE
# having more farms and being male leads to lower school enrollment
# supports my hypothesis of farm work affecting enrollment, since males might be expected to due more manual labor
maleedu_model <- allData %>%
  select(YEAR, pct_enrolled, n_farm, pct_male) %>%
  na.omit() %>%
  group_by(YEAR) %>%
  nest() %>%

  # Call map inside of mutate to add new columns to the tibble
  # upper and lower are the upper and lower bounds of the confidence interval created by tidy()

  # options
  #STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
         # acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_farm, horses, pop, male, female, 
        # enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate

  mutate(mod = map(data, ~ lm(pct_enrolled ~ n_farm*pct_male, data = .)),
         reg_results = map(mod, ~ tidy(., conf.int = TRUE)),
         n_farm_coef = map_dbl(reg_results, ~ filter(., term == "n_farm") %>% 
                                   pull(estimate)),
         n_farm_upper = map_dbl(reg_results, ~ filter(., term == "n_farm") %>% 
                                    pull(conf.high)),
         n_farm_lower = map_dbl(reg_results, ~ filter(., term == "n_farm") %>% 
                                    pull(conf.low)),
         pct_male_coef = map_dbl(reg_results, ~ filter(., term == "pct_male") %>% 
                                   pull(estimate)),
         pct_male_upper = map_dbl(reg_results, ~ filter(., term == "pct_male") %>% 
                                    pull(conf.high)),
         pct_male_lower = map_dbl(reg_results, ~ filter(., term == "pct_male") %>% 
                                    pull(conf.low)),
         interaction_coef = map_dbl(reg_results, ~ filter(., term == "n_farm:pct_male") %>% 
                                   pull(estimate)),
         interaction_upper = map_dbl(reg_results, ~ filter(., term == "n_farm:pct_male") %>% 
                                    pull(conf.high)),
         interaction_lower = map_dbl(reg_results, ~ filter(., term == "n_farm:pct_male") %>% 
                                    pull(conf.low))
         )
# Incorporate
maleedu_gt <- maleedu_model %>%
  ungroup() %>%

  # Select the columns needed, as per the instructions

  select(-data:-reg_results) %>%

  gt() %>%

  # Add the desired titles, subtitles, and column labels

  tab_header(title = "Effect of Gender and Number of Farms on Enrollment",
             subtitle = "With 95% Confidence Interval") %>%
  tab_spanner(label = "Farm Coefficients", columns = vars(n_farm_lower, n_farm_coef, n_farm_upper)) %>%
  tab_spanner(label = "Percent Male Coefficients", columns = vars(pct_male_lower, pct_male_coef,pct_male_upper)) %>%
  tab_spanner(label = "Interaction Coefficients", columns = vars(interaction_lower, interaction_coef, interaction_upper)) %>%
  cols_label(YEAR = "Year", n_farm_coef	= "Estimate", n_farm_upper = "Upper", n_farm_lower = "Lower", pct_male_coef = "Estimate", pct_male_upper = "Upper", pct_male_lower = "Lower",	interaction_coef = "Estimate", interaction_upper = "Upper", interaction_lower = "Lower") %>%
  # Round all values to two decimal places

  fmt_number(columns = vars(n_farm_lower, n_farm_coef, n_farm_upper, pct_male_lower, pct_male_coef,pct_male_upper, interaction_lower, interaction_coef, interaction_upper), decimals = 4)

# INCORPORATE
# Effect of gender on enrollment dwindles over time
maleedu_plot <- maleedu_model %>%
  ungroup() %>%
  select(YEAR, pct_male_coef, pct_male_upper, pct_male_lower) %>%
  mutate(YEAR = as.factor(YEAR)) %>%
  group_by(YEAR) %>%
  ggplot(aes(x = YEAR, y = pct_male_coef)) + 
  geom_errorbar(ymin = maleedu_model$pct_male_lower, ymax = maleedu_model$pct_male_upper) + 
  geom_point() +
  theme_classic() +
  ylim(-1.5, 0.5) + 
  labs(x = "Year",
       y = "Coefficient",
       title = "Percent Male Coefficients Over Time",
       subtitle = "Measuring the effect of being male on school enrollment")

# avg_value_implements and cows_per_farm have an effect on enrollment
# in general, for 1879 and 1890, more cows per farm = less enrollment
# aka bigger farms = less enrollment?
model2a <- modelData %>% 
  filter(YEAR == "1870") %>%
  lm(data = ., pct_enrolled ~ cows_per_farm) %>% 
  tidy(conf.int = TRUE) %>% 
  select(term, estimate, p.value, conf.low, conf.high) %>%
  gt() %>% 
  fmt_number(data = ., columns = c("estimate", "p.value", "conf.low", "conf.high"), 
             decimals = 2)

# there was an infinite value left in the cows_per_farm column for this year
# not sure why, but I got rid of it
model2a <- modelData %>% 
  filter(YEAR == "1890", !is.infinite(cows_per_farm)) %>%
  lm(data = ., pct_enrolled ~ cows_per_farm) %>% 
  tidy(conf.int = TRUE) %>% 
  select(term, estimate, p.value, conf.low, conf.high) %>%
  gt() %>% 
  fmt_number(data = ., columns = c("estimate", "p.value", "conf.low", "conf.high"), 
             decimals = 2)

model2c <- modelData %>% 
  filter(YEAR == "1910") %>%
  lm(data = ., pct_enrolled ~ cows_per_farm) %>% 
  tidy(conf.int = TRUE) %>% 
  select(term, estimate, p.value, conf.low, conf.high) %>%
  gt() %>% 
  fmt_number(data = ., columns = c("estimate", "p.value", "conf.low", "conf.high"), 
             decimals = 2)

model2
# cow plot
# incorporate
cowplot <- modelData %>% 
  filter(YEAR == "1870" | YEAR == "1890" | YEAR == "1910") %>%
  ggplot(aes(x = cows_per_farm, y = pct_enrolled, color = YEAR)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10(labels = comma) +
  labs(title = "Cows and Education",
       subtitle = "Relationship between the number of cows per farm\nand school enrollment, by county",
       x = "Average Cows per Farm",
       y = "Percent of Population Enrolled in School") +
  theme_classic()

modelData %>% 
  filter(YEAR == "1870" | YEAR == "1890" | YEAR == "1910") %>%
  ggplot(aes(x = cows_per_farm, y = pct_enrolled, color = YEAR)) +
  geom_point() +
  geom_smooth(se = FALSE)

# Enrollment/Illiteracy model (just for testing)

# This is saying that for every 1% increase in illiterate people, there is a 0.24% 
# decrease in school enrollment, which makes sense
model3 <- modelData %>%
  lm(data = ., pct_enrolled ~ pct_illiterate) %>%
  tidy(conf.int = TRUE) %>%
  select(term, estimate, p.value, conf.low, conf.high) %>%
  gt() %>%
  fmt_number(data = ., columns = c("estimate", "p.value", "conf.low", "conf.high"), 
             decimals = 2)

mod3plot <- modelData %>%
  filter(! is.na(pct_enrolled) & ! is.na(pct_illiterate)) %>%
  ggplot(aes(pct_enrolled, pct_illiterate)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

```{r moreModels}

# can only use percentages until adjusting for inflation
# OR deal with each year independently

# Set the NA values to almost zero
# Not ideal, but necessary for running regressions
# setting to zero returns NaNs

reg_years <- allData %>% 
  # sample the data to run faster
  # change this later
  sample_n(size = 5000) %>%
  # Get rid of the infinite cows_per_farm values
  filter(! (is.na(pct_enrolled) & is.na(avg_size))) %>%
  group_by(YEAR) %>%
  arrange(desc(YEAR)) %>% 
  nest()

  # Call map inside of mutate to add new columns to the tibble
  # reg_results are the results of the regression we ran in question 2
  # coef is the coefficient of log_migrants for each continent
  # upper and lower are the upper and lower bounds of the confidence interval created by tidy()

  # options
  #STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
         # acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_farm, horses, pop, male, female, 
        # enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate
reg_years1 <- reg_years %>%
  mutate(mod = map(data, ~ lm(pct_enrolled ~ avg_size + avg_value_implements + n_farm + pct_irrigated + cows_per_farm + pct_urban, data = .)),
         reg_results = map(mod, ~ tidy(., conf.int = TRUE)),
         avg_size_coef = map_dbl(reg_results, ~ filter(., term == "avg_size") %>% 
                                   pull(estimate)),
         avg_size_upper = map_dbl(reg_results, ~ filter(., term == "avg_size") %>% 
                                    pull(conf.high)),
         avg_size_lower = map_dbl(reg_results, ~ filter(., term == "avg_size") %>% 
                                    pull(conf.low)),
         avg_value_implements_coef = map_dbl(reg_results, ~ filter(., term == "avg_value_implements") %>% 
                                               pull(estimate)),
         avg_value_implements_upper = map_dbl(reg_results, ~ filter(., term == "avg_value_implements") %>% 
                                                pull(conf.high)),
         avg_value_implements_lower = map_dbl(reg_results, ~ filter(., term == "avg_value_implements") %>% 
                                                pull(conf.low)),
         n_farm_coef = map_dbl(reg_results, ~ filter(., term == "n_farm") %>% pull(estimate)),
         n_farm_upper = map_dbl(reg_results, ~ filter(., term == "n_farm") %>% pull(conf.high)),
         n_farm_lower = map_dbl(reg_results, ~ filter(., term == "n_farm") %>% pull(conf.low)),
         pct_irrigated_coef = map_dbl(reg_results, ~ filter(., term == "pct_irrigated") %>% pull(estimate)),
         pct_irrigated_upper = map_dbl(reg_results, ~ filter(., term == "pct_irrigated") %>% pull(conf.high)),
         pct_irrigated_lower = map_dbl(reg_results, ~ filter(., term == "pct_irrigated") %>% pull(conf.low)),
         cows_per_farm_coef = map_dbl(reg_results, ~ filter(., term == "cows_per_farm") %>% pull(estimate)),
         cows_per_farm_upper = map_dbl(reg_results, ~ filter(., term == "cows_per_farm") %>% pull(conf.high)),
         cows_per_farm_lower = map_dbl(reg_results, ~ filter(., term == "cows_per_farm") %>% pull(conf.low)),
         pct_urban_coef = map_dbl(reg_results, ~ filter(., term == "pct_urban") %>% pull(estimate)),
         pct_urban_upper = map_dbl(reg_results, ~ filter(., term == "pct_urban") %>% pull(conf.high)),
         pct_urban_lower = map_dbl(reg_results, ~ filter(., term == "pct_urban") %>% pull(conf.low))
  )

# Now use the resulting tibble to create the gt table

q4_gt <- q4 %>%

  # The original year column was giving me problems because it's a grouping variable
  # Ungrouping solved the problem

  ungroup() %>%

  # Select the columns needed, as per the instructions

  select(YEAR, coef, lower, upper) %>%

  gt() %>%

  # Add the desired titles, subtitles, and column labels

  tab_header(title = "Title",
             subtitle = "With 95% Confidence Interval") %>%
  cols_label(YEAR = "Year", coef = "Migrant Coefficient", 
             lower = "Lower Bound", upper = "Upper Bound") %>%

  # Round all values to two decimal places

  fmt_number(columns = vars(coef, lower, upper), decimals = 2)

```

### SHINY FINAL
```{r shiny, eval=FALSE}

# PASTE DATA and LIBRARIES here after

# Input

ui <- fluidPage(

    # flatly is a nice theme to make the tabs look put together

    # Application title

    navbarPage("Agriculture and Education in the U.S.", theme = shinytheme("flatly"),

               # remember that tab panel is for what shows up in the tab, and title panel is 
               # for what shows up at the top of the page within the tab

               tabPanel("Tab 1",

                        # Give the page a title
                        titlePanel("Relationship Between Farm Value and Education, by County"),

                        # Generate a row with a sidebar
                        sidebarLayout(

                            # Define the sidebar with one input
                            # insteaad of region, I want year ######
                            sidebarPanel(
                                selectInput("year", "Year:", 
                                            choices=c("1870" = "1870", "1890" = "1890", "1910" = "1910"), 
                                            multiple = TRUE)),
                            helpText("From Great Plains Population and Environment Survey Agricultural and Social & Demographic Data.")
                        ),

                        # Create a spot for the plot
                        mainPanel(
                            tabsetPanel(id = "tabsMain",
                                        tabPanel("Plot",

                                                 # this tells the UI what plot to put

                                                 plotOutput("agPlot")  
                                        )))),
               tabPanel("Models",

                        # Give the page a title
                        titlePanel("Relationship Between Farm Value and Education, by County"),
                        p("Make some models! (See notes)"),
                        gt_output("model1")
               ),
               tabPanel("Maps",

                # Give the page a title
                titlePanel("Title Relating to Maps"),
                p("Add a slider for different years/variables & figure out how to get two maps side by side"),
                p("Maps may take up to 30 seconds to load."),
               leafletOutput("map1"),
               p()),

               tabPanel("About",

                        # Give the page a title
                        titlePanel("About this project"),
                        p("My name is Julia Englebert. I am a senior at Harvard College studying Near Eastern Languages and Civilizations with a secondary field in Ethnicity, Migration, and Rights. Some of my academic interests include the history and future of Midwestern agriculture, and the role of data science in shaping social policy."),
                        br(),
                        p("This project uses historical census and American Community Survey (ACS) data to explore correlations between agriculture and education in the Midwestern United States. By joining the two datasets, the former containing agricultural data and the later containing demographic data, I was able to look at the impact of farming, land value, crop value, and more on education in Midwestern communities over the course of a century. 
                        I chose this particular time range in part because of the data that was available to me, and in part due to the economic, industrial, and education policy changes that ocurred from 1870 to 1960. 
                          My data is sourced from the IPUMS National Historical Geographic Information System (NHGIS) at nhgis.org. IPUMS provides census data from around the world, but my project focuses specifically on the United States from 1870 to 1960. All datasets and gis files used for this project can be downloaded from the NHGIS Data Finder:"),
                        p(a(href="https://data2.nhgis.org/main", "NHGIS Data Finder")),
                        br(),
                        p("To see more of my work, please visit my Github account:"),
                        p(a(href="https://github.com/julia-englebert", "Julia Englebert")),
                        br(),
                        titlePanel("Citations"),
                        p("Steven Manson, Jonathan Schroeder, David Van Riper, and Steven Ruggles. IPUMS National Historical Geographic Information System: Version 14.0 [Database]. Minneapolis, MN: IPUMS. 2019. http://doi.org/10.18128/D050.V14.0"),
                        p("The College of William and Mary and the Minnesota Population Center. School Attendance Boundary Information System (SABINS): Version 1.0. Minneapolis, MN: University of Minnesota 2011.")
               )
    ))

# use plotly to allow hover text to appear on plots, giving more info about the point

server <- function(input, output) {

    # Fill in the spot we created for a plot
    output$agPlot <- renderPlot({

        # Render a plot
        b_1870_1910 %>% 
            filter(year %in% input$year) %>% 
            group_by(year) %>%
            ggplot(aes(x = value_farms, y = perc_school, color = year)) + 
            geom_point(alpha = 0.5) + 
            ylim(0, 100) + 
            geom_smooth(method = "lm") + 
            labs(title = "Relationship Between Farm Value and Education, by County", 
                 subtitle = "Great Plains Region, United States", 
                 x = "Total Dollar Value of Farms", 
                 y = "Percent Students Ages 5-24 Enrolled in School") +
            scale_x_log10(labels = comma) +
            theme_classic()
    })

    output$map1 <- renderLeaflet({
            data1870 <- data1870j %>%
                filter(STATE == "North Dakota" | STATE == "South Dakota" | STATE == "Nebraska" | STATE == "Minnesota" | STATE == "Iowa" | STATE == "Missouri" | STATE == "Wisconsin" | STATE == "Illinois" | STATE == "Kansas" | STATE == "Michigan" | STATE == "Indiana" | STATE == "Ohio", AJX001 > 0) %>%
                mutate(valueAllFarmProds1870 = AJX001) %>%
                select(GISJOIN, valueAllFarmProds1870)

            # Import Census Tract Shapefile into R as SpatialPolygonsDataFrameFormat (SP Dataframe)
            # dsn is location of folder which contains shapefiles, (.proj, .shp etc.)
            # layer is the filename of the .shp file inside the
            # folder dsn points to.

            county1870 <- sf::st_read(dsn = "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1870",
                                      layer = "US_county_1870_conflated")

            county1870 <-
                county1870 %>%
                merge(data1870, "GISJOIN")

            # Set projection of tracts dataset to `projection` required by leaflet

            county1870 <- sf::st_transform(county1870, crs="+init=epsg:4326")

            # Condense size of data for faster processing

            county1870 <- rmapshaper::ms_simplify(county1870)

            # Set palette color

            pal <- colorNumeric("viridis", NULL)

            #  Plot the data

            leaflet(county1870) %>%
                addTiles() %>%
                addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
                            fillColor = ~pal(valueAllFarmProds1870)) %>%
                addLegend(pal = pal, values = ~valueAllFarmProds1870, opacity = 1.0, title = "1870 Value of All Farm Productions,\nby Dollar")
    })

    output$model1 <- render_gt({

        # experimental model

        model1

    })
}

# Run the application

shinyApp(ui = ui, server = server)
```

# DO NOT MAKE CHANGES BELOW THIS LINE

### Map Code

```{r genericGis}

# DO NOT CHANGE THIS

# might also want to add a filter for state??
# if time permits...

# gis template for any year/variable desired
# only change these two lines!
year <- 1870
myvar <- "Average Farm Value"

datayear <- mwData %>%
  # use get() to use the stored variable
  # idk why, but it doesn't work without this
  # the variable must be converted to a double in the original mwData dataset
  mutate(myvariable = (get(myvar))) %>%
  filter(YEAR == year, myvariable > 0) %>%
  select(GISJOIN, myvariable)

# Import Census Tract Shapefile into R as SpatialPolygonsDataFrameFormat (SP Dataframe)
# dsn is location of folder which contains shapefiles, (.proj, .shp etc.)
# layer is the filename of the .shp file inside the
# folder dsn points to.

dsnyear <- shapefiles %>% 
  filter(Year == year) %>% 
  pull(Dsn)
layeryear <- shapefiles %>% 
  filter(Year == year) %>% 
  pull(Layer)

countyyear <- sf::st_read(dsn = dsnyear,
                         layer = layeryear)

countyyear <-
  countyyear %>%
  merge(datayear, "GISJOIN")

# Set projection of tracts dataset to `projection` required by leaflet

countyyear <- sf::st_transform(countyyear, crs="+init=epsg:4326")

# Condense size of data for faster processing

countyyear <- rmapshaper::ms_simplify(countyyear)

# Set palette color

pal <- colorNumeric("viridis", NULL)

#  Plot the data

# use shiny to add a title, which will be be equivalent to myvar

leaflet(countyyear) %>%
  addTiles() %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
              fillColor = ~pal(myvariable)) %>%
  addLegend(pal = pal, values = ~myvariable, opacity = 1.0, title = myvar)

```

# Save This
```{r shiny, eval=FALSE}

# Pasting this here in case my computer explodes

# This is a Shiny web application. You can run the application by clicking
# the 'Run App' button above.
#
# Find out more about building applications with Shiny here:
#
#    http://shiny.rstudio.com/
#

library(shiny)
library(haven)
library(tidyverse)
library(gt)
library(devtools)
library(tibble)
library(foreign)
library(scales)
library(shinythemes)
library(plotly)
library(leaflet)
library(broom)

# packages for mapping
library(leaflet)
library(spdplyr)
library(rmapshaper)
library(sf)

# The real data

data1870a <- read.csv("raw-data/nhgis0002_csv/nhgis0002_ds16_1870_county.csv")
data1870b <- read.csv("raw-data/nhgis0002_csv/nhgis0002_ds17_1870_county.csv")
data1870j <- full_join(data1870a, data1870b, by = c("GISJOIN", "STATE", "COUNTY"))

data1880a <- read.csv("raw-data/nhgis0002_csv/nhgis0002_ds22_1880_county.csv") 
data1880b <- read.csv("raw-data/nhgis0002_csv/nhgis0002_ds23_1880_county.csv")
data1880j <- full_join(data1880a, data1880b, by = c("GISJOIN", "STATE", "COUNTY"))

data1890a <- read.csv("raw-data/nhgis0002_csv/nhgis0002_ds26_1890_county.csv")
data1890b <- read.csv("raw-data/nhgis0002_csv/nhgis0002_ds27_1890_county.csv")
data1890j <- full_join(data1890a, data1890b, by = c("GISJOIN", "STATE", "COUNTY"))

data1800s <- bind_rows(data1870j, data1880j, data1890j)

# I'm not actually using this data anymore
# I just wanted to make sure I could get my plots running smoothly
tbl_1 <- read_por("ICPSR_04254_ag/DS0001/04254-0001-Data.por", col_select = c(NAME, UNFIPS, FML_XX_A, ARE_XX_A, FRM_XX_Q, PRP_XX_V, IMP_XX_V, AFP_XX_V)) %>%
    rename(name = NAME, fips = UNFIPS, land_in_farms = FML_XX_A, total_area = ARE_XX_A, num_farms = FRM_XX_Q, value_farms = PRP_XX_V, value_implements = IMP_XX_V, value_production_betterments_stocks = AFP_XX_V)

tbl_3 <- read_por("ICPSR_04254_ag/DS0003/04254-0003-Data.por", col_select = c(NAME, UNFIPS, FML_XX_A, ARE_XX_A, FRM_XX_Q, PRP_XX_V, IMP_XX_V, AFP_XX_V))%>%
    rename(name = NAME, fips = UNFIPS, land_in_farms = FML_XX_A, total_area = ARE_XX_A, num_farms = FRM_XX_Q, value_farms = PRP_XX_V, value_implements = IMP_XX_V, value_production_betterments_stocks = AFP_XX_V)

tbl_5 <- read_por("ICPSR_04254_ag/DS0005/04254-0005-Data.por", col_select = c(NAME, UNFIPS, FML_XX_A, ARE_XX_A, FRM_XX_Q, PRP_OF_V, IMP_XX_V, CRO_XX_V)) %>%
    rename(name = NAME, fips = UNFIPS, land_in_farms = FML_XX_A, total_area = ARE_XX_A, num_farms = FRM_XX_Q, value_implements = IMP_XX_V, value_crops = CRO_XX_V) %>%
    mutate(value_farms = PRP_OF_V) %>%
    mutate(value_livestock = "NA", value_productsNotFed = "NA") %>%
    select(name, fips, land_in_farms, total_area, num_farms, value_farms, value_implements, value_livestock, value_productsNotFed)

tbl_1_d <- read_por("ICPSR_04296_demog/DS0001/04296-0001-Data.por", col_select = c(NAME, UNFIPS, TTOT, FT0_4, FT5_9, FT10_14, FT15_19, FT20_24, FT25_29, MT0_4, MT5_9, MT10_14, MT15_19, MT20_24, MT25_29, POP10P, SCH)) %>%
    rename(name = NAME, fips = UNFIPS, total = TTOT, people_over_10 = POP10P, pupils_in_school = SCH) %>%
    mutate(fem_5_24 = FT5_9 + FT10_14 + FT15_19 + FT20_24, mal_5_24 = MT5_9 + MT10_14 + MT15_19 + MT20_24)

tbl_3_d <- read_por("ICPSR_04296_demog/DS0003/04296-0003-Data.por", col_select = c(NAME, UNFIPS, TTOT, FT0_4, FT5_9, FT10_14, FT15_19, FT20_24, FT25_29, MT0_4, MT5_9, MT10_14, MT15_19, MT20_24, MT25_29, SCH)) %>%
    rename(name = NAME, fips = UNFIPS, total = TTOT, pupils_in_school = SCH) %>%
    mutate(fem_5_24 = FT5_9 + FT10_14 + FT15_19 + FT20_24, mal_5_24 = MT5_9 + MT10_14 + MT15_19 + MT20_24)

tbl_5_d <- read_por("ICPSR_04296_demog/DS0005/04296-0005-Data.por", col_select = c(NAME, UNFIPS, TTOT, FT0_4, FT5_9, FT10_14, FT15_19, FT20_24, FT25_29, MT0_4, MT5_9, MT10_14, MT15_19, MT20_24, MT25_29, SCH69, SCH1014, SCH1517, SCH1820)) %>% 
    rename(name = NAME, fips = UNFIPS, total = TTOT) %>%
    mutate(fem_5_24 = FT5_9 + FT10_14 + FT15_19 + FT20_24, mal_5_24 = MT5_9 + MT10_14 + MT15_19 + MT20_24,
           # Inconsistent age ranges
           # Not sure how to fix this
           pupils_in_school_6_20 = SCH69 + SCH1014 + SCH1517 + SCH1820)

j_1870 <- full_join(tbl_1, tbl_1_d, by = "fips") %>% 
    mutate(perc_school = pupils_in_school/(fem_5_24 + mal_5_24)*100) %>%
    select(name.x, fips, value_farms, perc_school) %>%
    group_by(fips)

j_1870 <- na.omit(j_1870)

j_1890 <- full_join(tbl_3, tbl_3_d, by = "fips") %>%
    mutate(perc_school = pupils_in_school/(fem_5_24 + mal_5_24)*100) %>%
    select(name.x, fips, value_farms, perc_school) %>%
    group_by(fips)

j_1890 <- na.omit(j_1890)

j_1910 <- full_join(tbl_5, tbl_5_d, by = "fips") %>%
    mutate(perc_school = pupils_in_school_6_20/(fem_5_24 + mal_5_24)*100) %>%
    select(name.x, fips, value_farms, perc_school) %>%
    group_by(fips)

j_1910 <- na.omit(j_1910)

b_1870_1910 <- bind_rows("1870" = j_1870, "1890" = j_1890, "1910" = j_1910, .id = "year") %>%
    group_by(year)

# this should be done by state, not fips!!!
model1 <- lm(data = b_1870_1910, perc_school ~ fips + value_farms) %>% 
    tidy(conf.int = TRUE) %>% 
    gt() %>%
    fmt_number(data = ., columns = c("estimate", "std.error", "statistic", "p.value", "conf.low", "conf.high"), decimals = 2)

# Input

ui <- fluidPage(

    # flatly is a nice theme to make the tabs look put together

    # Application title

    navbarPage("Agriculture and Education in the U.S.", theme = shinytheme("flatly"),

               # remember that tab panel is for what shows up in the tab, and title panel is 
               # for what shows up at the top of the page within the tab

               tabPanel("Tab 1",

                        # Give the page a title
                        titlePanel("Relationship Between Farm Value and Education, by County"),

                        # Generate a row with a sidebar
                        sidebarLayout(

                            # Define the sidebar with one input
                            # insteaad of region, I want year ######
                            sidebarPanel(
                                selectInput("year", "Year:", 
                                            choices=c("1870" = "1870", "1890" = "1890", "1910" = "1910"), 
                                            multiple = TRUE)),
                            helpText("From Great Plains Population and Environment Survey Agricultural and Social & Demographic Data.")
                        ),

                        # Create a spot for the plot
                        mainPanel(
                            tabsetPanel(id = "tabsMain",
                                        tabPanel("Plot",

                                                 # this tells the UI what plot to put

                                                 plotOutput("agPlot")  
                                        )))),
               tabPanel("Models",

                        # Give the page a title
                        titlePanel("Relationship Between Farm Value and Education, by County"),
                        p("Make some models! (See notes)"),
                        gt_output("model1")
               ),
               tabPanel("Maps",

                # Give the page a title
                titlePanel("Title Relating to Maps"),
                p("Add a slider for different years/variables & figure out how to get two maps side by side"),
                p("Maps may take up to 30 seconds to load."),
               leafletOutput("map1"),
               p()),

               tabPanel("About",

                        # Give the page a title
                        titlePanel("About this project"),
                        p("My name is Julia Englebert. I am a senior at Harvard College studying Near Eastern Languages and Civilizations with a secondary field in Ethnicity, Migration, and Rights. Some of my academic interests include the history and future of Midwestern agriculture, and the role of data science in shaping social policy."),
                        br(),
                        p("This project uses historical census and American Community Survey (ACS) data to explore correlations between agriculture and education in the Midwestern United States. By joining the two datasets, the former containing agricultural data and the later containing demographic data, I was able to look at the impact of farming, land value, crop value, and more on education in Midwestern communities over the course of a century. 
                        I chose this particular time range in part because of the data that was available to me, and in part due to the economic, industrial, and education policy changes that ocurred from 1870 to 1960. 
                          My data is sourced from the IPUMS National Historical Geographic Information System (NHGIS) at nhgis.org. IPUMS provides census data from around the world, but my project focuses specifically on the United States from 1870 to 1960. All datasets and gis files used for this project can be downloaded from the NHGIS Data Finder:"),
                        p(a(href="https://data2.nhgis.org/main", "NHGIS Data Finder")),
                        br(),
                        p("To see more of my work, please visit my Github account:"),
                        p(a(href="https://github.com/julia-englebert", "Julia Englebert")),
                        br(),
                        titlePanel("Citations"),
                        p("Steven Manson, Jonathan Schroeder, David Van Riper, and Steven Ruggles. IPUMS National Historical Geographic Information System: Version 14.0 [Database]. Minneapolis, MN: IPUMS. 2019. http://doi.org/10.18128/D050.V14.0"),
                        p("The College of William and Mary and the Minnesota Population Center. School Attendance Boundary Information System (SABINS): Version 1.0. Minneapolis, MN: University of Minnesota 2011.")
               )
    ))

# use plotly to allow hover text to appear on plots, giving more info about the point

server <- function(input, output) {

    # Fill in the spot we created for a plot
    output$agPlot <- renderPlot({

        # Render a plot
        b_1870_1910 %>% 
            filter(year %in% input$year) %>% 
            group_by(year) %>%
            ggplot(aes(x = value_farms, y = perc_school, color = year)) + 
            geom_point(alpha = 0.5) + 
            ylim(0, 100) + 
            geom_smooth(method = "lm") + 
            labs(title = "Relationship Between Farm Value and Education, by County", 
                 subtitle = "Great Plains Region, United States", 
                 x = "Total Dollar Value of Farms", 
                 y = "Percent Students Ages 5-24 Enrolled in School") +
            scale_x_log10(labels = comma) +
            theme_classic()
    })

    output$map1 <- renderLeaflet({
            data1870 <- data1870j %>%
                filter(STATE == "North Dakota" | STATE == "South Dakota" | STATE == "Nebraska" | STATE == "Minnesota" | STATE == "Iowa" | STATE == "Missouri" | STATE == "Wisconsin" | STATE == "Illinois" | STATE == "Kansas" | STATE == "Michigan" | STATE == "Indiana" | STATE == "Ohio", AJX001 > 0) %>%
                mutate(valueAllFarmProds1870 = AJX001) %>%
                select(GISJOIN, valueAllFarmProds1870)

            # Import Census Tract Shapefile into R as SpatialPolygonsDataFrameFormat (SP Dataframe)
            # dsn is location of folder which contains shapefiles, (.proj, .shp etc.)
            # layer is the filename of the .shp file inside the
            # folder dsn points to.

            county1870 <- sf::st_read(dsn = "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1870",
                                      layer = "US_county_1870_conflated")

            county1870 <-
                county1870 %>%
                merge(data1870, "GISJOIN")

            # Set projection of tracts dataset to `projection` required by leaflet

            county1870 <- sf::st_transform(county1870, crs="+init=epsg:4326")

            # Condense size of data for faster processing

            county1870 <- rmapshaper::ms_simplify(county1870)

            # Set palette color

            pal <- colorNumeric("viridis", NULL)

            #  Plot the data

            leaflet(county1870) %>%
                addTiles() %>%
                addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
                            fillColor = ~pal(valueAllFarmProds1870)) %>%
                addLegend(pal = pal, values = ~valueAllFarmProds1870, opacity = 1.0, title = "1870 Value of All Farm Productions,\nby Dollar")
    })

    output$model1 <- render_gt({

        # experimental model

        model1

    })
}

# Run the application

shinyApp(ui = ui, server = server)

```